Assessing environmental and predator impacts on Antarctic Krill (*Euphausia superba*) population dynamics from an integrated length-to-age assessment model perspective

### 2025-08-29 MS2_Krill.Rmd
adding last chgs

---
title: "Assessing environmental and predator impacts on Antarctic Krill (*Euphausia superba*) population dynamics from an integrated length-to-age assessment model perspective"
author:
  - Mauricio Mardones ^[Universidad de Magallanes, mamardon@umag.cl; Instituto Milenio Base, Chile; Centro IDEAL, Chile; Instituto Español de Oceanografía, Spain]
  - Erica Jarvis Mason^[Antarctic División, SouthWest NOAA, La Jolla, California, USA]
  - Andrea Pinones ^[Universidad Austral de Chile; Centro IDEAL, Chile; Instituto Milenio Base, Chile]
  - Lucas Krüger ^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Chile; Instituto Milenio Base, Chile]
  - César Cárdenas^[Instituto Antártico Chileno; Instituto Milenio Base, Chile]
  - Richard Methot ^[Fishery Science División, NorthWeast NOAA, Seattle, USA]
output:
  bookdown::pdf_document2:
    reference_docx: MS2_Krill_template.docx.docx
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
csl: apa.csl
#csl: icesjournal.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.5
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---

@@

Understanding the population dynamics of Antarctic krill (*Euphausia superba*) is critical for achieving sustainable and ecosystem-based management fisheries in the Western Antarctic Peninsula. In this study, we implemented an integrated length-to-age model that not only incorporates diverse data sources related to krill population dynamics, but also enables the assessment of ecosystem influences—specifically environmental variability and predation pressure—on recruitment, biomass, and stock status. We explored four scenarios that either accounted for or ignored ecosystem influences on krill population dynamics, yielding results consistent with the expected behavior of a stock assessment framework tailored to this type of species. Preliminary performance analyses reveal systematic differences across scenarios, highlighting increased bias in estimates of fishing mortality (F) and spawning stock biomass (SSB) in models that include ecosystem variables. Conversely, the inclusion of environmental and predator data results in lower estimates of recruitment, total biomass, and spawning biomass, offering a more precautionary—and arguably more realistic—representation of krill population dynamics. Despite ongoing challenges in incorporating ecosystem variables into integrated models like this, we argue that such factors should be embedded directly into the state equations that govern population dynamics, rather than addressed through pre and post analyses. Overlooking these drivers may lead to less accurate conclusions, hence limiting the effectiveness of management recommendations. Integrated modeling approaches that draw on extensive biological, ecological, fishery, and environmental datasets provide a robust foundation for developing management strategies aligned with CCAMLR’s objectives for conservation and sustainable use. Such approaches are particularly relevant for ecologically complex systems like the krill fishery in the Western Antarctic Peninsula.

@@

Increasing evidence demonstrates that environmental conditions—such as sea surface temperature, sea ice extent, and phytoplankton availability—strongly influence krill recruitment, survival, and spatial distribution [@Murphy2007; @Flores2012; @Saba2014; @Morley2020; @Perry2020; @Walsh2020; @Vorrath2020; @Veytia2021]. Similarly, fluctuations in predator abundance and distribution can exert significant top-down control on krill populations [@Moosa2017; @Reisinger2022; @Baines2022; @Riaz2023; @Biuw2024]. There is now broad scientific consensus on the need for stock assessment models that incorporate multiple data sources, capture temporal and spatial variability, and explicitly account for ecosystem interactions [@Marshall2019]. Neglecting these ecosystem drivers can lead to biased estimates of stock status and suboptimal management advice [@Haltuch2009; @Richards2016; @Marshall2019; @Crone2019]. Therefore, stock assessment frameworks aligned with CCAMLR’s new strategy should explicitly incorporate these ecosystem components to more accurately represent krill population processes.

Integrating ecological complexity into krill stock assessments presents multiple challenges. Long-term data collection is often affected by changes in monitoring programs and sampling methodologies [@Hill2019], while some key regions of the WAP lack sufficient data to construct reliable krill population indices. Additional difficulties arise from indirect drivers—such as environmental variability and predator-prey interactions—which are difficult to quantify due to spatial and temporal heterogeneity, non-linear responses, and the combined effects of multiple stressors [@Constable2023]. Despite these challenges, substantial data exist on krill and associated environments. Fishery monitoring programs provide detailed biological data from catches through onboard scientific observers. Ecosystem monitoring, like the CCAMLR Ecosystem Monitoring Program (CEMP), tracks predator populations and environmental variability at key sites. Together, these initiatives support ecosystem-based management by offering insights into both direct fishery impacts and the broader ecological context influencing krill dynamics.

This diversity of data creates an opportunity to develop integrated assessment frameworks that account for the coupled biological, ecological, and fishery processes shaping krill populations. Spatially explicit approaches include the State-Space Assessment Model (SAM) [@Nielsen2021] and the Globally Applicable Area-Disaggregated General Ecosystem Toolbox (Gadget) [@Bartolino2011]. For statistical catch-at-age assessments, widely used frameworks include the C++ Algorithmic Stock Assessment Laboratory (CASAL) [@Bull2004] and Stock Synthesis (SS3) [@Methot2013], among others. These advanced tools offer robust methodologies to address existing data gaps while incorporating the full complexity of ecosystem and population dynamics [@Hoyle2020].

This study aims to integrate various data sources on krill (*Euphausia superba*), including fishery operations (e.g. catches and effort distribution), biological characteristics (e.g. growth rates and reproductive parameters), spatiotemporal population dynamics, and key ecosystem drivers (e.g. environmental forcing and predator-prey interactions) into a unified stock assessment framework. Specifically, we use an integrated length-to-age model implemented in Stock Synthesis (SS3), which enables the explicit inclusion of environmental covariates affecting biomass, recruitment patterns, and productivity, as well as the potential impacts of predator populations in natural mortality configuration. This approach strengthens the scientific foundation for sustainable management recommendations, ensuring that krill harvesting strategies align with CCAMLR's ecosystem-based management approach [@Constable2011; @Cavanagh2016; @Hill2019; @Constable2023; @Zhao2023; @CCAMLRWGSAM2024; @WGEMM2024]. The primary objective is to evaluate the statistical performance of alternative model configurations representing different hypotheses about ecosystem influences using a suite of diagnostic tools and performance metrics commonly applied in contemporary stock assessment practices. Our results provide empirical evidence for the value of ecosystem-informed model-based approach, demonstrating that the inclusion of environmental and predator covariates has impact in model fit, predictive skill, and structural uncertainty.

@@

@@

>>>>

---
title: "Assessing environmental and predator impacts on Antarctic Krill (*Euphausia superba*) population dynamics from an integrated length-to-age assessment model perspective"
subtitle: "Version 2"
author:
  - Mauricio Mardones ^[Universidad de Magallanes, mamardon@umag.cl; Instituto Milenio Base, Chile; Centro IDEAL, Chile; Instituto Español de Oceanografía, Spain]
  - Erica Jarvis Mason^[Antarctic División, SouthWest NOAA, La Jolla, California, USA]
  - Andrea Pinones ^[Universidad Austral de Chile; Centro IDEAL, Chile; Instituto Milenio Base, Chile]
  - Lucas Krüger ^[Instituto Antártico Chileno, Instituto Milenio Base, Chile]
  - Francisco Santa Cruz ^[Instituto Antártico Chileno, Universidad de Magallanes, Chile; Instituto Milenio Base, Chile]
  - César Cárdenas^[Instituto Antártico Chileno; Instituto Milenio Base, Chile]
  - Richard Methot ^[Fishery Science División, NorthWeast NOAA, Seattle, USA]
output:
  bookdown::word_document2:
    reference_docx: MS2_Krill_template.docx.docx
    number_sections: false
fig_caption: yes
bibliography: SA_krill.bib
csl: apa.csl
#csl: icesjournal.csl
link-citations: yes
toc: false
linkcolor: blue
linestretch: 1.5
header-includes:
- \fontsize{12}{16}
- \selectfont
- \usepackage{lscape}
---

@@

Understanding the population dynamics of Antarctic krill (*Euphausia superba*) is critical for achieving sustainable and ecosystem-based management fisheries in the Western Antarctic Peninsula. In this study, we implemented an integrated length-to-age model that not only incorporates diverse data sources related to krill population dynamics, but also enables the assessment of ecosystem influences—specifically environmental variability and predation pressure—on recruitment, biomass, and stock status. We explored four scenarios that either accounted for or ignored ecosystem influences on krill population dynamics, yielding results consistent with the expected behavior of a stock assessment framework tailored to this type of species. Preliminary performance analyses reveal systematic differences across scenarios, highlighting increased bias in estimates of fishing mortality (F) and spawning stock biomass (SSB) in models that include ecosystem variables. Conversely, the inclusion of environmental and predator data results in lower estimates of recruitment, total biomass, and spawning biomass, offering a more precautionary, and arguably more realistic, representation of krill population dynamics. Despite ongoing challenges in incorporating ecosystem variables into integrated models like this, we argue that such factors should be embedded directly into the state equations that govern population dynamics, rather than addressing them in peripheral analyses. Overlooking these drivers may lead to less accurate conclusions, hence limiting the effectiveness of management recommendations. Integrated modeling approaches that draw on extensive biological, ecological, fishery, and environmental datasets provide a robust foundation for developing management strategies aligned with CCAMLR’s objectives for conservation and sustainable use. Such integrative approaches provide a stronger foundation for CCAMLR's objectives, particularly in managing the ecologically complex krill fishery of the Western Antarctic Peninsula.

@@

Increasing evidence demonstrates that environmental conditions, such as sea surface temperature, sea ice extent, and phytoplankton availability, strongly influence krill recruitment, survival, and spatial distribution [@Murphy2007; @Flores2012; @Saba2014; @Morley2020; @Perry2020; @Walsh2020; @Vorrath2020; @Veytia2021]. Similarly, fluctuations in predator abundance and distribution can exert significant top-down control on krill populations [@Moosa2017; @Reisinger2022; @Baines2022; @Riaz2023; @Biuw2024]. There is now broad scientific consensus on the need for stock assessment models that incorporate multiple data sources, capture temporal and spatial variability, and explicitly account for ecosystem interactions [@Marshall2019]. Neglecting these ecosystem drivers can lead to biased estimates of stock status and suboptimal management advice [@Haltuch2009; @Richards2016; @Marshall2019; @Crone2019; @Robertson2024]. Therefore, stock assessment frameworks aligned with CCAMLR’s new strategy should explicitly incorporate these ecosystem components to more accurately represent krill population processes.

Integrating ecological complexity into krill stock assessments presents multiple challenges. Long-term data collection is often affected by changes in monitoring programs and sampling methodologies [@Hill2019], while some key regions of the WAP lack sufficient data to construct reliable krill population indices. Additional difficulties arise from indirect drivers, such as environmental variability and predator, prey interactions—which are difficult to quantify due to spatial and temporal heterogeneity, non-linear responses, and the combined effects of multiple stressors [@Constable2023]. Despite these challenges, substantial data exist on krill and associated environments. Fishery monitoring programs provide detailed biological data from catches through onboard scientific observers. Ecosystem monitoring, like the CCAMLR Ecosystem Monitoring Program (CEMP), tracks predator populations and environmental variability at key sites. Together, these initiatives support ecosystem-based management by offering insights into both direct fishery impacts and the broader ecological context influencing krill dynamics.

This diversity of data creates an opportunity to develop integrated assessment frameworks that account for the coupled biological, ecological, and fishery processes shaping krill populations, which have traditionally been assessed from a data-limited context using model-based approaches. A wide variety of integrated models allow the use of diverse data sources, such as spatially explicit approaches including the State-Space Assessment Model (SAM) [@Nielsen2021] and the Globally Applicable Area-Disaggregated General Ecosystem Toolbox (Gadget) [@Bartolino2011]. For statistical catch-at-age assessments, widely used frameworks include the C++ Algorithmic Stock Assessment Laboratory (CASAL) [@Bull2004] and Stock Synthesis (SS3) [@Methot2013], among others. These advanced tools offer robust methodologies to address existing data gaps while incorporating the full complexity of ecosystem and population dynamics [@Hoyle2020].

This study aims to integrate various data sources on krill (Euphausia superba), including fishery operations (e.g. catches and effort distribution), biological characteristics (e.g. growth rates and reproductive parameters), spatiotemporal population dynamics, and key ecosystem drivers (e.g. environmental forcing and predator-prey interactions) into a unified stock assessment framework. Specifically, we use an integrated length-to-age model, which enables the explicit inclusion of environmental covariates affecting biomass, recruitment patterns, and productivity, as well as the potential impacts of predator populations in natural mortality configuration. The primary objective is to evaluate the performance of alternative model configurations that represent different hypotheses about ecosystem influences, using diagnostic tools and performance metrics commonly applied in contemporary stock assessments. Consistent with CCAMLR’s ecosystem-based management framework [@Constable2011; @Cavanagh2016; @Hill2019; @Constable2023; @Zhao2023; @CCAMLRWGSAM2024; @WGEMM2024], this approach provides empirical evidence for the value of ecosystem-informed models, showing that incorporating environmental and predator covariates has an impact on model fit, predictive skill, and structural uncertainty, and this has direct implications for sustainable management recommendations and precautionary krill harvesting strategies.

@@

# Discussion

This research represents an advance the understanding Antarctic krill (*Euphausia superba*) population dynamics by addressing two fundamental objectives. First, it consolidates and harmonizes multiple long-term datasets collected over decades, particularly from Subarea 48.1 in the Western Antarctic Peninsula, into a unified statistical framework, responding to a recurrent management priority about intregrated model in krill emphasized by the CCAMLR. Second, it integrates ecosystem variables, such as environmental drivers and predator impacts, thereby moving beyond the traditional single-stocks species approach and taking an initial step toward ecosystem-informed stock assessments. This approach not only enhances the biological realism of krill population modelling but also reveals the structural complexity and statistical challenges that must be addressed to ensure its effective application in management Antarctic fishery contexts.

### Model Performance and Diagnostic Evaluation

Increasing model complexity can destabilize internal matrices and processes, potentially leading to convergence issues [@Kinzey2015; @Kinzey2019; @Dong2025]. While all models in this exercise successfully converged and produced Hessian matrices suitable for quantifying parameter uncertainty, the comprehensive diagnostic evaluation revealed important trade-offs between biological and ecological realism and statistical performance. While the ecosystem-informed models (particularly `s1.4`) showed improved biological realism through reduced Pearson residuals and better representation of length-frequency distributions, they also exhibited increased retrospective bias and reduced likelihood values compared to the baseline model (`s1.1`). The elevated retrospective bias in fishing mortality estimates for ecosystem scenarios, particularly the pronounced bias (rho = 6.912) in scenario `s1.4`, suggests potential overfitting or structural instability when both environmental and predator variables are simultaneously incorporated. The observed bias patterns may also reflect temporal mismatches between ecosystem variables and population responses, suggesting that lagged effects or non-linear relationships should be explored in future model development.

The shift toward ecosystem-based fisheries management highlights the need for stock assessment models that explicitly incorporate species interactions, environmental drivers, and predator pressures, as addressed in this analysis. For Antarctic krill, integrating ecosystem variables—such as predation pressure and environmental variability—can improve the realism of population dynamics; however, it also increases model complexity. This, in turn, affects model fit and raises parameter uncertainty, revealing a trade-off between traditional single-species assessments and ecosystem-informed approaches, as described by @Collie2016 and @Robertson2024. This chapter embraced that challenge and confirmed the inherent difficulties, yet we argue that this approach is a step in the right direction. It offers valuable insights for the evolving management strategy that CCAMLR seeks to implement for the krill fishery.

### Spatial Considerations and the Areas-as-Fleets Approach

Building on the analyses from [Chapter 3](#chap3), recognizing and accounting for spatial heterogeneity in krill populations across the large area of Subarea 48.1 was essential, particularly for determining how to incorporate this dimension into the integrated model. To address this, we applied the areas-as-fleets approach, a methodology that captures spatial differences and incorporates them into global estimates of population variables. This framework allows for a more realistic representation of local variability in abundance, selectivity, and population processes [@Nielsen2021; @Waterhouse2014]. However, some Management Units (MUs), which served as the structural basis for the areas-as-fleets approach, had limited data—for example, Joinville Island, with few years of survey and mfihery monitoring data. While these data limitations were not explicitly evaluated in this thesis, they could introduce complications in future integrated krill modeling efforts and should be carefully considered in subsequent analyses.

### Environmental Drivers and Recruitment Dynamics

The identification of chlorophyll-a concentration as the primary environmental driver affecting krill recruitment (see [Chapter 2](#chap2)) highlights the central role of primary productivity in shaping krill population dynamics [@Ainley2017; @Constable2023; @Saba2014].  This finding is consistent with recent studies emphasizing that krill recruitment is primarily driven by environmental conditions rather than spawning stock biomass [@ichii2023; @Perry2020; @Veytia2021].  The research identified chlorophyll-a concentration (Chla) as the most influential environmental predictor for krill recruitment. Models incorporating environmental variables (`s1.3` and `s1.4`) showed significantly lower Akaike Information Criterion (AIC) values, indicating better model performance and parsimony compared to models without environmental data. While these results highlight that integrating such environmental variables introduces complexity to model structure and performance, it also enhances model realism, a key challenge currently faced by stock assessment approaches. By explicitly integrating variables that are mechanistically linked to key population processes—such as recruitment—these models move beyond purely stochastic representations of stock productivity [@Haltuch2009; @Crone2019]. Such integration allows for a more process informed understanding of population variability, improving both explanatory power and predictive capacity.

### Predator-Prey Interactions and Natural Mortality

The integration of penguin predation data into natural mortality estimates represents a novel approach in an integrated stock assessment, and even more so for krill. This size-selective predation pattern is biologically consistent with foraging theory and observed predator behavior, where no significant differences were identified amongkey predators like *Pygoscelis* penguins in the size ranges of krill in their diet [@Panasiuk2024]. The declining trend in the synoptic predator biomass index, particularly after 2015, reflects documented population declines in Antarctic Peninsula penguin colonies [@Emmerson2022; @Kruger2019; @Salmeron2023; @Trivelpiece2011a]. These declines, primarily affecting Adélie and Chinstrap penguin populations, have been attributed to climate change impacts, sea ice loss, and shifts in prey availability [@Kruger2019]. The incorporation of this declining predation pressure into the stock assessment model provides a mechanism for capturing ecosystem mediated changes in natural mortality that would otherwise be overlooked in traditional single-species approaches.

Regarding scenarios of modeling, the scenarios with explicit inclusion of predator components (`s1.2` and `s1.4`), showed increased bias in estimates of fishing mortality (F) and spawning stock biomass (SSB), they resulted in lower estimates of recruitment, total biomass, and spawning biomass. Incorporating predation mortality into stock assessment models is essential for accurately representing population dynamics in marine ecosystems and offering a more precautionary and arguably more realistic representation of krill dynamics. Efforts to incorporate predation as a modelling component in marine stock assessment processes are not new. @Hollowed2000 revealed higher total natural mortality rates than traditional life-history estimates and showed that ignoring predation uncertainty underestimated stock biomass uncertainty by up to 20%. Similarly, @Richards2016 introduced a predation pressure index for northern shrimp, combining predator biomass and diet frequency to scale natural mortality over time. This method improved model fit and reduced retrospective bias without requiring absolute predator consumption data, which are often unavailable. For Antarctic krill stock assessments, incorporating predation variability improves mortality estimates and supports ecosystem-based management.

### Implications for Precautionary Management

While there are concerns about the performance of models that consider ecosystem variables, their estimates of key population variables—including spawning biomass, total biomass, and recruitment—are generally lower than those produced by models without ecosystem variables (S1).  The significant reduction in spawning biomass estimates between scenarios `s1.2` and `s1.4` (mean difference: 2.6 million tonnes) demonstrates that neglecting ecosystem variables may lead to overly optimistic stock assessments and potentially unsustainable harvest recommendations.  This conservative tendency could have significant implications for fisheries management, as these more precautionary estimates align with CCAMLR's new strategic guidelines, which call for the integration of ecosystem variables into predictive models to support precautionary management approaches. The more conservative estimates produced by ecosystem-informed models provide a buffer against uncertainty in population status, which is particularly important given the high variability and unpredictability of Antarctic marine ecosystems [@Hill2016]. However, the increased model complexity and associated statistical challenges must be balanced against the improved biological realism to ensure that management advice remains scientifically credible and operationally feasible.

### Model Limitations and Future Directions

Several limitations of the current modeling approach warrant discussion. First, the integration of predator effects relied on relatively coarse scale abundance indices that may not fully capture fine-scale predator-prey interactions or behavioral responses. This is primarily due to two reasons: the large spatial scale of Area 48.1, which covers more than 658,739 $km^{2}$, and the fact that CCAMLR's ecosystem observer deployment does not have observation points distributed throughout this entire area, which hampers comprehensive data collection across the region. In this sense, other predators such as whales [@Pallin2023; @Reisinger2022], which have a much greater impact on krill availability and abundance, were not included in this analysis. This exclusion was due to the high variability in methods and estimates used to assess whale abundances and consumption rates across time series, particularly in the Western Antarctic Peninsula (WAP), which complicates the integration of these data in a uniform framework. Studies suggest that whale consumption rates may even exceed around 20 times the total krill catch [@Biuw2024], highlighting the necessity of incorporating these key predators in future work to more accurately represent predation mortality. More detailed predator foraging data, including spatial overlap indices and prey switching behavior, could further enhance this analysis. Second, the environmental covariate framework, while effective in capturing chlorophyll driven recruitment variability, represents a simplified view of the complex environmental drivers affecting krill populations. Incorporating multivariate ecosystem variables must also be coupled with the development of models capable of handling such complexity; otherwise, contradictory signals may arise, potentially reducing model performance and compromising the reliability of outputs. Future studies should consider multivariate environmental relationships, threshold effects, and lagged responses that may better represent the mechanistic links between climate variability and krill population dynamics. Third, the current model assumes static relationships between ecosystem variables and parameters estimations. However, climate change may alter these relationships over time, requiring adaptive modeling approaches that can accommodate non-stationary ecosystem conditions. The development of time-varying parameter frameworks or regime specific models may be necessary to address this challenge.

Despite these limitations, and considering that this is an academic-scientific work, we strongly believe that this represents the future direction for the development of integrated krill models. Such approaches provide a foundation to generate scientifically grounded recommendations to CCAMLR, employing the most advanced fisheries science methods currently used worldwide.

@@

---

This research demonstrates that integrating ecosystem variables into krill stock assessment models provides important advances in both biological realism and precautionary management advice. By harmonizing diverse long-term datasets and embedding environmental and predator covariates within an integrated framework, the work highlights both the potential and the challenges of moving beyond traditional single-species assessments. The findings show that ecosystem-informed models have a measurable impact on estimates of recruitment, mortality, and biomass, often producing more conservative outcomes that align with CCAMLR’s precautionary management principles. At the same time, the increased model complexity introduces statistical trade-offs and uncertainty that require careful evaluation and transparent communication.

Looking forward, future efforts should refine ecosystem integration by improving predator-prey datasets, exploring multivariate and non-linear environmental relationships, and adopting adaptive modelling frameworks capable of capturing non-stationary dynamics under climate change. Despite the inherent challenges, this research supports the view that ecosystem-informed assessment is both necessary and feasible, offering a pathway for CCAMLR to strengthen its ecosystem-based management of the Antarctic krill fishery.

---
